---
title: "SST_patterns"
author: "Amieroh Abrahams"
date: "19 August 2019"
output: html_document
---

To accurately compare 
  Start date CMC start date 1991-09-02 end date 2017-03-15
  Start date OISST start date 1981-09- 01  and end date 2018-09-28
  Start date SACTN start date 1973 and end date 2017-12-16
  
```{r}
#glm to determine if there is a difference in the number of events recorded at different at every distance

load("Data/Count/OISST_count.RData")
load("Data/Count/G1SST_count.RData")
load("Data/Count/MUR_count.RData")
load("Data/Count/CMC_count.RData")


options(scipen = 999)
anova_func <- function(df){
  sites_aov <- aov(distance ~ duration * site * site, data = df)
return(sites_aov)
}

OISST_anov <- anova_func(df = OISST_upwell_base)
summary(OISST_anov)
CMC_anov <- anova_func(df = CMC_upwell_base)
summary(CMC_anov)

MUR_anov <- anova_func(df = MUR_upwell_base)
summary(MUR_anov)

SACTN_anov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
summary(SACTN_anov)




lm_coeff <- function(df){
  res <- lm(formula = val ~ date_peak, data = df)
  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
}
# Changes in upwelling metrics
lm_func <- function(df){
  upwell_lm <- df %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var, distance) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
}

OISST_lm <- lm_func(df = OISST_upwell_base)
CMC_lm <- lm_func(df = CMC_upwell_base)

SACTN_lm <- SACTN_upwell_base %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
```
  

As a common date we will use a start date from 1992-01-31 and end date 16-12-31

```{r}
# Generalosed linear models

# Generalized Linear Models- Look at the difference in metrics of in situ upwelling and remotely sensed SST data upwelling 
# Generalized linear models are fit using the glm( ) function. The form of the glm function is glm(formula, family=familytype(link=linkfunction), data=)

# trends_duration_func <- function(df){
#   upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
#                      family = poisson(link = "log"))
#   upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
# }
# 
# OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
# CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
# SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
# 
# 
# ### Intensity: Chnaged the family and link because of the negative values within the dataset 
# trends_intensity_func <- function(df){
#   upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
#                      family = gaussian(link = "identity"))
#   upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
# }
# 
# OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
# CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
# SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
```

# From the years 1992 - 2017

```{r}
# Which pixels showed the highest upwelling counts within the different datasets
# Which site showed the most intense upwelling within the different datasets

metric_func <- function(df){
 metrics <-df %>% 
  filter(year(date_start) %in% 1992:2017) %>% 
  #mutate(year = year(date_start)) %>% 
  group_by(distance, site) %>% 
  summarise(mean_intensity = mean(intensity_mean),
            sum_events = sum(event_no)) %>% 
   select(site, distance, mean_intensity, sum_events)
}

OISST_metrics <- metric_func(OISST_upwell_base)
CMC_metrics <- metric_func(CMC_upwell_base)

# No of u pwelling events found at the different sites
SACTN_metrics <- SACTN_upwell_base %>% 
  filter(year(date_start) %in% 1992:2017) %>% 
  group_by(site) %>% 
  summarise(mean_intensity = mean(intensity_mean),
            sum_events = sum(event_no)) %>% 
  select(site, mean_intensity, sum_events)

# Create a table showing these metrics
```

# Wind plots

```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 15}
library(ggradar)
library(dplyr)
library(scales)
library(tibble)
load("Data/UI_angle.RData")

# Detects NA values, no NA values are present
# Plotting wind variables
source("Functions/wind.rose.R") # This it may be related to the countmax = NA in the function

wind_daily_renamed <- UI_angle %>% 
  mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>% 
  dplyr::rename(spd = mean_speed) %>%
  dplyr::rename(dir = dir_circ) %>% 
  filter(spd > 0)

p.wr2 <- plot.windrose(data = wind_daily_renamed,
              spd = "spd",
              dir = "dir")

p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
p.wr3

################3

# load("Data/UI_angle.RData")
# 
# wind_plot_prep <- UI_angle %>% 
#   select(-t, -ui.saws)
# 
# wind_plot <- wind_plot_prep %>% 
#   as_tibble(rownames = "site") %>% 
#   group_by(site) %>% 
#   mutate_at(vars(-site), rescale) %>% 
#   tail(4) 
# 
# ggradar(wind_plot)
# 
# ??countmax
```

# ANOVA
- Relationship between duration, year and site as a function of the distance 
The number of events per distance and then compare this at each distance 

```{r}
options(scipen = 999)
anova_func <- function(df){
  sites_aov <- aov(distance ~ duration * site * site, data = df)
return(sites_aov)
}

OISST_anov <- anova_func(df = OISST_upwell_base)
summary(OISST_anov)
CMC_anov <- anova_func(df = CMC_upwell_base)
summary(CMC_anov)

MUR_anov <- anova_func(df = MUR_upwell_base)
summary(MUR_anov)

SACTN_anov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
summary(SACTN_anov)
```


Plot showing upwelling at the different sites for each of the different distances. Scatter plot

```{r}

load("Data/OISST_upwell_base.RData")
load("Data/CMC_upwell_base.RData")
load("Data/SACTN_upwell_base.RData")
load("Data/MUR_upwell_base.RData")
load("Data/G1SST_upwell_base.RData")

km_func <- function(df){
  upwell_base <-  df%>% 
      mutate(distance_km = case_when(distance == "10000" ~ "10",
                                    distance == "20000" ~ "20",
                                    distance  == "30000" ~"30",
                                    distance  == "40000" ~ "40",
                                    distance  == "50000" ~ "50"))
}
OISST_upwell_base <- km_func(df = OISST_upwell_base)
OISST_upwell_base$distance_km <- as.numeric(OISST_upwell_base$distance_km)

CMC_upwell_base <- km_func(df = CMC_upwell_base)
CMC_upwell_base$distance_km <- as.numeric(CMC_upwell_base$distance_km)

MUR_upwell_base <- km_func(df = MUR_upwell_base)
MUR_upwell_base$distance_km <- as.numeric(MUR_upwell_base$distance_km)

G1SST_upwell_base <- km_func(df = G1SST_upwell_base)
G1SST_upwell_base$distance_km <- as.numeric(G1SST_upwell_base$distance_km)

library(ggpubr)
year_func <- function(df){
  Filtered <- df %>%
filter(year(date_start) %in% seq(2011, 2013, 1))
}

CMC_filtered <- year_func(df = CMC_upwell_base)
OISST_filtered <- year_func(df = OISST_upwell_base)
SACTN_filtered <- year_func(df = SACTN_upwell_base)
MUR_filtered <- year_func(df = MUR_upwell_base)
G1SST_filtered <- year_func(df = G1SST_upwell_base)

CMC_SEAPOINT <- CMC_filtered %>% 
  filter(site == "Sea Point")
CMC_SB <- CMC_filtered %>% 
  filter(site == "Saldanha Bay")
CMC_PN <- CMC_filtered %>% 
  filter(site == "Port Nolloth")
CMC_LB <- CMC_filtered %>% 
  filter(site == "Lamberts Bay")

###################
## OISST
OISST_SEAPOINT <- OISST_filtered %>% 
  filter(site == "Sea Point")
OISST_SB <- OISST_filtered %>% 
  filter(site == "Saldanha Bay")
OISST_PN <- OISST_filtered %>% 
  filter(site == "Port Nolloth")
OISST_LB <- OISST_filtered %>% 
  filter(site == "Lamberts Bay")

############
#####MUR

MUR_SEAPOINT <- MUR_filtered %>% 
  filter(site == "Sea Point")
MUR_SB <- MUR_filtered %>% 
  filter(site == "Saldanha Bay")
MUR_PN <- MUR_filtered %>% 
  filter(site == "Port Nolloth")
MUR_LB <- MUR_filtered %>% 
  filter(site == "Lamberts Bay")
###########
####G1SST

G1SST_SEAPOINT <- G1SST_filtered %>% 
  filter(site == "Sea Point")
G1SST_SB <- G1SST_filtered %>% 
  filter(site == "Saldanha Bay")
G1SST_PN <- G1SST_filtered %>% 
  filter(site == "Port Nolloth")
G1SST_LB <- G1SST_filtered %>% 
  filter(site == "Lamberts Bay")

```

# Lolli plot

```{r}
plot_loli_func <- function(df){
  ggplot(df, aes(x = date_peak, y = duration)) + 
  geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
  #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
  facet_wrap(~distance_km, ncol = 2) +
  scale_y_continuous(breaks = c(5,10,15,20,25,30),
                     limits = c(1,30)) +
    scale_x_date(limits = as.Date(c("2011-01-01","2014-01-01"))) +
  theme_bw() +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=17)) +
  theme(strip.text.x = element_text(size = 17))+
  theme(panel.spacing = unit(1, "lines")) +
    #theme(strip.text = c("10km", "20km", "30km", "40km", "50km")) +
  #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
  xlab("Date (years)") + ylab("Event duration (days)") 
}


(OISST_SP_pl <- plot_loli_func(df = OISST_SEAPOINT))
(OISST_SB_pl <- plot_loli_func(df = OISST_SB))
(OISST_PN_pl <- plot_loli_func(df = OISST_PN))
(OISST_LB_pl <- plot_loli_func(df = OISST_LB))

OISST_SP_pl <- OISST_SP_pl +
  ggtitle("D.") #Sea Point
OISST_SB_pl <- OISST_SB_pl +
  ggtitle("C.") # Saldanha Bay
OISST_PN_pl <- OISST_PN_pl +
  ggtitle("B.")        # Port Nolloth
OISST_LB_pl <- OISST_LB_pl +
  ggtitle("A.")#Lamberts Bay

combined_OISST_pl_2 <- ggarrange(OISST_SB_pl, OISST_SP_pl) 

combined_OISST_pl_1 <- ggarrange(OISST_LB_pl, OISST_PN_pl)

#######
# CMC

(CMC_SP_pl <- plot_loli_func(df = CMC_SEAPOINT))
(CMC_SB_pl <- plot_loli_func(df = CMC_SB))
(CMC_PN_pl <- plot_loli_func(df = CMC_PN))
(CMC_LB_pl <- plot_loli_func(df = CMC_LB))

CMC_SP_pl <- CMC_SP_pl +
  ggtitle("D.") #Sea Point
CMC_SB_pl <- CMC_SB_pl +
  ggtitle("C.") # Saldanha Bay
CMC_PN_pl <- CMC_PN_pl +
  ggtitle("B.")        # Port Nolloth
CMC_LB_pl <- CMC_LB_pl +
  ggtitle("A.")#Lamberts Bay

combined_CMC_pl_2 <- ggarrange(CMC_SB_pl, CMC_SP_pl) 

combined_CMC_pl_1 <- ggarrange(CMC_LB_pl, CMC_PN_pl)

##########
### MUR
(MUR_SP_pl <- plot_loli_func(df = MUR_SEAPOINT))
(MUR_SB_pl <- plot_loli_func(df = MUR_SB))
(MUR_PN_pl <- plot_loli_func(df = MUR_PN))
(MUR_LB_pl <- plot_loli_func(df = MUR_LB))


MUR_SP_pl <- MUR_SP_pl +
  ggtitle("D.") #Sea Point
MUR_SB_pl <- MUR_SB_pl +
  ggtitle("C.") # Saldanha Bay
MUR_PN_pl <- MUR_PN_pl +
  ggtitle("B.")        # Port Nolloth
MUR_LB_pl <- MUR_LB_pl +
  ggtitle("A.")#Lamberts Bay


combined_MUR_pl_2 <- ggarrange(MUR_SB_pl, MUR_SP_pl) 

combined_MUR_pl_1 <- ggarrange(MUR_LB_pl, MUR_PN_pl)
#######
#G1SST

(G1SST_SP_pl <- plot_loli_func(df = G1SST_SEAPOINT))
(G1SST_SB_pl <- plot_loli_func(df = G1SST_SB))
(G1SST_PN_pl <- plot_loli_func(df = G1SST_PN))
(G1SST_LB_pl <- plot_loli_func(df = G1SST_LB))

G1SST_SP_pl <- G1SST_SP_pl +
  ggtitle("D.") #Sea Point
G1SST_SB_pl <- G1SST_SB_pl +
  ggtitle("C.") # Saldanha Bay
G1SST_PN_pl <- G1SST_PN_pl +
  ggtitle("B.")        # Port Nolloth
G1SST_LB_pl <- G1SST_LB_pl +
  ggtitle("A.")#Lamberts Bay

combined_G1SST_pl_2 <- ggarrange(G1SST_SB_pl, G1SST_SP_pl) 

combined_G1SST_pl_1 <- ggarrange(G1SST_LB_pl, G1SST_PN_pl)

########

plot_loli_func <- function(df){
  ggplot(df, aes(x = date_peak, y = duration)) + 
  geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
  #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
   facet_wrap(~site, ncol = 2) +
  scale_y_continuous(breaks = c(5,10,15,20,25,30),
                     limits = c(1,30)) +
  theme_bw() +
  theme(axis.text=element_text(size=17),
        axis.title=element_text(size=17)) +
  theme(strip.text.x = element_text(size = 17))+
  theme(panel.spacing = unit(0.5, "lines")) +
  #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
  xlab("Date (years)") + ylab("Event duration (days)") 
}

(SACTN_plot <- plot_loli_func(df = SACTN_filtered))
```

Loading the data
    This is the data obtained once the upwelling signals were filtered out

```{r}
# load("Data/OISST_upwell_base.RData")
# load("Data/CMC_upwell_base.RData")
# load("Data/SACTN_upwell_base.RData")
# load("Data/MUR_upwell_base.RData")
# load("Data/G1SST_upwell_base.RData")

load("Data_coast_base/OISST_upwell_base.RData")
load("Data_coast_base/CMC_upwell_base.RData")
load("Data_coast_base/SACTN_upwell_base.RData")
load("Data_coast_base/MUR_upwell_base.RData")
load("Data_coast_base/G1SST_upwell_base.RData")

# Removing the distance of 20 and 40kms
library(dplyr)
removing_distance_func <- function(df){
  removing_dist_func<- df %>% 
       filter(!distance %in% c(20000,40000))
}

OISST_final <- removing_distance_func(df = OISST_upwell_base)
G1SST_final <- removing_distance_func(df = G1SST_upwell_base)
MUR_final <- removing_distance_func(df = MUR_upwell_base)
CMC_final <- removing_distance_func(df = CMC_upwell_base)

# save(OISST_final, file = "Data/OISST_final.RData")
# save(G1SST_final, file = "Data/G1SST_final.RData")
# save(MUR_final, file = "Data/MUR_final.RData")
# save(CMC_final, file = "Data/CMC_final.RData")
```

To accurately compare 
  Start date CMC start date 1991-09-02 end date 2017-03-15
  Start date OISST start date 1981-09- 01  and end date 2018-09-28
  Start date SACTN start date 1973 and end date 2017-12-16
  
```{r}
#glm to determine if there is a difference in the number of events recorded at different at every distance

load("Data/Count/OISST_count.RData")
load("Data/Count/G1SST_count.RData")
load("Data/Count/MUR_count.RData")
load("Data/Count/CMC_count.RData")


options(scipen = 999)
anova_func <- function(df){
  sites_aov <- aov(distance ~ duration * site * site, data = df)
return(sites_aov)
}

OISST_anov <- anova_func(df = OISST_upwell_base)
summary(OISST_anov)
CMC_anov <- anova_func(df = CMC_upwell_base)
summary(CMC_anov)

MUR_anov <- anova_func(df = MUR_upwell_base)
summary(MUR_anov)

SACTN_anov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
summary(SACTN_anov)




lm_coeff <- function(df){
  res <- lm(formula = val ~ date_peak, data = df)
  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
}
# Changes in upwelling metrics
lm_func <- function(df){
  upwell_lm <- df %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var, distance) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
}

OISST_lm <- lm_func(df = OISST_upwell_base)
CMC_lm <- lm_func(df = CMC_upwell_base)

SACTN_lm <- SACTN_upwell_base %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
```
  

# ANOVA
- Relationship between duration, year and site as a function of the distance 
The number of events per distance and then compare this at each distance 

```{r}
options(scipen = 999)
anova_func <- function(df){
  sites_aov <- aov(distance ~ duration * site * site, data = df)
return(sites_aov)
}

OISST_anov <- anova_func(df = OISST_upwell_base)
summary(OISST_anov)
CMC_anov <- anova_func(df = CMC_upwell_base)
summary(CMC_anov)

MUR_anov <- anova_func(df = MUR_upwell_base)
summary(MUR_anov)

SACTN_anov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
summary(SACTN_anov)
```


Plot showing upwelling at the different sites for each of the different distances. Scatter plot

```{r}

load("Data/OISST_upwell_base.RData")
load("Data/CMC_upwell_base.RData")
load("Data/SACTN_upwell_base.RData")
load("Data/MUR_upwell_base.RData")
load("Data/G1SST_upwell_base.RData")

km_func <- function(df){
  upwell_base <-  df%>% 
      mutate(distance_km = case_when(distance == "10000" ~ "10",
                                    distance == "20000" ~ "20",
                                    distance  == "30000" ~"30",
                                    distance  == "40000" ~ "40",
                                    distance  == "50000" ~ "50"))
}
OISST_upwell_base <- km_func(df = OISST_upwell_base)
OISST_upwell_base$distance_km <- as.numeric(OISST_upwell_base$distance_km)

CMC_upwell_base <- km_func(df = CMC_upwell_base)
CMC_upwell_base$distance_km <- as.numeric(CMC_upwell_base$distance_km)

MUR_upwell_base <- km_func(df = MUR_upwell_base)
MUR_upwell_base$distance_km <- as.numeric(MUR_upwell_base$distance_km)

G1SST_upwell_base <- km_func(df = G1SST_upwell_base)
G1SST_upwell_base$distance_km <- as.numeric(G1SST_upwell_base$distance_km)

library(ggpubr)
year_func <- function(df){
  Filtered <- df %>%
filter(year(date_start) %in% seq(2011, 2013, 1))
}

CMC_filtered <- year_func(df = CMC_upwell_base)
OISST_filtered <- year_func(df = OISST_upwell_base)
SACTN_filtered <- year_func(df = SACTN_upwell_base)
MUR_filtered <- year_func(df = MUR_upwell_base)
G1SST_filtered <- year_func(df = G1SST_upwell_base)

CMC_SEAPOINT <- CMC_filtered %>% 
  filter(site == "Sea Point")
CMC_SB <- CMC_filtered %>% 
  filter(site == "Saldanha Bay")
CMC_PN <- CMC_filtered %>% 
  filter(site == "Port Nolloth")
CMC_LB <- CMC_filtered %>% 
  filter(site == "Lamberts Bay")

###################
## OISST
OISST_SEAPOINT <- OISST_filtered %>% 
  filter(site == "Sea Point")
OISST_SB <- OISST_filtered %>% 
  filter(site == "Saldanha Bay")
OISST_PN <- OISST_filtered %>% 
  filter(site == "Port Nolloth")
OISST_LB <- OISST_filtered %>% 
  filter(site == "Lamberts Bay")

############
#####MUR

MUR_SEAPOINT <- MUR_filtered %>% 
  filter(site == "Sea Point")
MUR_SB <- MUR_filtered %>% 
  filter(site == "Saldanha Bay")
MUR_PN <- MUR_filtered %>% 
  filter(site == "Port Nolloth")
MUR_LB <- MUR_filtered %>% 
  filter(site == "Lamberts Bay")
###########
####G1SST

G1SST_SEAPOINT <- G1SST_filtered %>% 
  filter(site == "Sea Point")
G1SST_SB <- G1SST_filtered %>% 
  filter(site == "Saldanha Bay")
G1SST_PN <- G1SST_filtered %>% 
  filter(site == "Port Nolloth")
G1SST_LB <- G1SST_filtered %>% 
  filter(site == "Lamberts Bay")

```

# Lolli plot

```{r}
plot_loli_func <- function(df){
  ggplot(df, aes(x = date_peak, y = duration)) + 
  geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
  #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
  facet_wrap(~distance_km, ncol = 2) +
  scale_y_continuous(breaks = c(5,10,15,20,25,30),
                     limits = c(1,30)) +
    scale_x_date(limits = as.Date(c("2011-01-01","2014-01-01"))) +
  theme_bw() +
  theme(axis.text=element_text(size=15),
        axis.title=element_text(size=17)) +
  theme(strip.text.x = element_text(size = 17))+
  theme(panel.spacing = unit(1, "lines")) +
    #theme(strip.text = c("10km", "20km", "30km", "40km", "50km")) +
  #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
  xlab("Date (years)") + ylab("Event duration (days)") 
}


(OISST_SP_pl <- plot_loli_func(df = OISST_SEAPOINT))
(OISST_SB_pl <- plot_loli_func(df = OISST_SB))
(OISST_PN_pl <- plot_loli_func(df = OISST_PN))
(OISST_LB_pl <- plot_loli_func(df = OISST_LB))

OISST_SP_pl <- OISST_SP_pl +
  ggtitle("D.") #Sea Point
OISST_SB_pl <- OISST_SB_pl +
  ggtitle("C.") # Saldanha Bay
OISST_PN_pl <- OISST_PN_pl +
  ggtitle("B.")        # Port Nolloth
OISST_LB_pl <- OISST_LB_pl +
  ggtitle("A.")#Lamberts Bay

combined_OISST_pl_2 <- ggarrange(OISST_SB_pl, OISST_SP_pl) 

combined_OISST_pl_1 <- ggarrange(OISST_LB_pl, OISST_PN_pl)

#######
# CMC

(CMC_SP_pl <- plot_loli_func(df = CMC_SEAPOINT))
(CMC_SB_pl <- plot_loli_func(df = CMC_SB))
(CMC_PN_pl <- plot_loli_func(df = CMC_PN))
(CMC_LB_pl <- plot_loli_func(df = CMC_LB))

CMC_SP_pl <- CMC_SP_pl +
  ggtitle("D.") #Sea Point
CMC_SB_pl <- CMC_SB_pl +
  ggtitle("C.") # Saldanha Bay
CMC_PN_pl <- CMC_PN_pl +
  ggtitle("B.")        # Port Nolloth
CMC_LB_pl <- CMC_LB_pl +
  ggtitle("A.")#Lamberts Bay

combined_CMC_pl_2 <- ggarrange(CMC_SB_pl, CMC_SP_pl) 

combined_CMC_pl_1 <- ggarrange(CMC_LB_pl, CMC_PN_pl)

##########
### MUR
(MUR_SP_pl <- plot_loli_func(df = MUR_SEAPOINT))
(MUR_SB_pl <- plot_loli_func(df = MUR_SB))
(MUR_PN_pl <- plot_loli_func(df = MUR_PN))
(MUR_LB_pl <- plot_loli_func(df = MUR_LB))


MUR_SP_pl <- MUR_SP_pl +
  ggtitle("D.") #Sea Point
MUR_SB_pl <- MUR_SB_pl +
  ggtitle("C.") # Saldanha Bay
MUR_PN_pl <- MUR_PN_pl +
  ggtitle("B.")        # Port Nolloth
MUR_LB_pl <- MUR_LB_pl +
  ggtitle("A.")#Lamberts Bay


combined_MUR_pl_2 <- ggarrange(MUR_SB_pl, MUR_SP_pl) 

combined_MUR_pl_1 <- ggarrange(MUR_LB_pl, MUR_PN_pl)
#######
#G1SST

(G1SST_SP_pl <- plot_loli_func(df = G1SST_SEAPOINT))
(G1SST_SB_pl <- plot_loli_func(df = G1SST_SB))
(G1SST_PN_pl <- plot_loli_func(df = G1SST_PN))
(G1SST_LB_pl <- plot_loli_func(df = G1SST_LB))

G1SST_SP_pl <- G1SST_SP_pl +
  ggtitle("D.") #Sea Point
G1SST_SB_pl <- G1SST_SB_pl +
  ggtitle("C.") # Saldanha Bay
G1SST_PN_pl <- G1SST_PN_pl +
  ggtitle("B.")        # Port Nolloth
G1SST_LB_pl <- G1SST_LB_pl +
  ggtitle("A.")#Lamberts Bay

combined_G1SST_pl_2 <- ggarrange(G1SST_SB_pl, G1SST_SP_pl) 

combined_G1SST_pl_1 <- ggarrange(G1SST_LB_pl, G1SST_PN_pl)

########

plot_loli_func <- function(df){
  ggplot(df, aes(x = date_peak, y = duration)) + 
  geom_lolli(colour = "steelblue3", colour_n = "navy", n = 3) + 
  #scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
   facet_wrap(~site, ncol = 2) +
  scale_y_continuous(breaks = c(5,10,15,20,25,30),
                     limits = c(1,30)) +
  theme_bw() +
  theme(axis.text=element_text(size=17),
        axis.title=element_text(size=17)) +
  theme(strip.text.x = element_text(size = 17))+
  theme(panel.spacing = unit(0.5, "lines")) +
  #strip=strip.custom(var.name=c("10km", "20km", "30km", "40km", "50km")) +
  xlab("Date (years)") + ylab("Event duration (days)") 
}

(SACTN_plot <- plot_loli_func(df = SACTN_filtered))
```

# GLM - Determining if a change occur in the number of signals detected at different distances from the coastline
# Included a table in the paper
  Table discuss/count the number of upwelling signals detected at the different distances
  
  - For some of the sites it is seen that fewer upwelling signals are visible at a further distance from the coastline
  - When adding up all the signals observed at their respective distance it is seen that there are only slight differences in the amount of signals detected In some cases the signals detected at 10km from the coastline exceed that detected at 50km and visa versa

## Number of events detected

Using the start date and getting the month and year into a column. The month and year will be on the xaxis
Then count the number of times per month at different years signals are detected

```{r}
load("Data/OISST_final.RData")
load("Data/CMC_final.RData")
load("Data/SACTN_upwell_base.RData")
load("Data/MUR_final.RData")
load("Data/G1SST_final.RData")

month_year_func <- function(df){
  month_year_func<- df %>% 
    mutate(month = month(date_start, abbr = T, label = T)) %>% 
    separate(date_start, into = c("year", "month_no")) %>% 
    unite(month_year, month,year)
    # Now need to get the year into a seperate column from the start date. So now date_start is split in the dataset
}

OISST_prep <- month_year_func(df = OISST_final)
CMC_prep <- month_year_func(df = CMC_final)
G1SST_prep <- month_year_func(df = G1SST_final)
MUR_prep <- month_year_func(df = MUR_final)

# Combined dataset
# Combined_frequency <- read_excel("Data/Combined_frequency.xlsx")
# save(Combined_frequency, file = "Data/Combined_frequency.RData")
```

# Plotting

```{r}
load("Data/Combined_frequency.RData")

Combined_frequency <- Combined_frequency %>%
  mutate(t = parse_date(date, format = "%b_%Y"),
         distance = factor(distance))

ggplot(data = Combined_frequency, aes(x = distance, y = number_of_signals)) +
  geom_boxplot(aes(colour = product)) +
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
  #geom_smooth(aes(colour = product), method = "lm") +
  #labs(x = NULL) +
  facet_wrap(~site) 


ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals, colour = product)) +
geom_point(aes(size = distance)) +
geom_smooth(method = "lm", size = 1.2)+
theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))

# Does not accurately show the months
# Plot need to include month_year, number of signals, distance and product ()

ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
  geom_line(aes(colour = product)) +
  geom_smooth(aes(colour = product), method = "lm") +
  labs(x = NULL) +
  facet_wrap(~site) #+
  # theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))


ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
  geom_line(aes(colour = distance)) +
  geom_point(aes(colour = distance)) +
  geom_smooth(aes(colour = distance), method = "lm") +
  # labs(x = NULL) +
  facet_wrap(~site) #+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
  
  ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
  geom_line(aes(colour = distance)) +
  geom_point(aes(colour = distance)) +
  geom_smooth(aes(colour = distance), method = "lm") +
  # labs(x = NULL) +
  facet_wrap(~site) #+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))

####

combined_frequency_split <- Combined_frequency %>% 
  separate(date, into = c("month", "year"))

# ggplot(data = combined_frequency_split, aes(x = year, y = number_of_signals)) +
#   geom_line(aes(colour = month)) +
#   geom_smooth(aes(colour = month), method = "lm") +
#   facet_wrap(~site)
```

#Pairwise correlations
Comparing the relaionship of variables within the same sample

  -Correlate the 1st pixel with the second pixel and the first pixel with the last pixel
  -Correlation between satellite data and number of signals at the different distances from the coastline
  
```{r}

# chi squared test

load("Data/Combined_frequency.RData")

combined_freq_spread <- pivot_wider(Combined_frequency, names_prefix = "dist_",
                                    names_from = distance, values_from = number_of_signals)


slope_calc <- function(df){
  df %>% 
    # mutate(row_num = 1:n()) %>% 
    # do(mod1 = lm(number_of_signals ~ row_num, data = .),
       # mod2 = lm(distance ~ row_num, data = .),
       # mod3 = lm(distance ~ number_of_signals, data = .),
       # mod4 = cor(.$distance, .$number_of_signals, method = "pearson", use = "complete.obs")[1]) %>% 
        # mutate(distance_slope = summary(mod1)$coeff[2],
        #    signal_slope = summary(mod2)$coeff[2],
        #    signal_distance_slope = summary(mod3)$coeff[2],
        #    signal_distance_r = mod4[1],
        #    signal_distance_r2 = glance(mod3)$adj.r.squared) %>%
        # select(-mod1, -mod2, -mod3, -mod4) %>% 
    do(mod1 = cor(.$dist_10, .$dist_30, method = "pearson", use = "complete.obs"),
       mod2 = cor(.$dist_10, .$dist_50, method = "pearson", use = "complete.obs")) %>%
    mutate(dist_10_vs_30_r = mod1[1],
           dist_10_vs_50_r = mod2[1]) %>% 
    # select(dist_10_vs_30_r, dist_10_vs_50_r) %>% 
    select(-mod1, -mod2) %>% 
    mutate_if(is.numeric, round, 2)
}

# test <- cor(combined_freq_spread$`10`, combined_freq_spread$`30`)

distance_corr <- combined_freq_spread %>% 
  group_by(site, product) %>% 
  slope_calc()




```


```{r}
library(hrbrthemes)
library(viridis)

Combined_frequency <- Combined_frequency %>%
  mutate(t = parse_date(date, format = "%b_%Y"),
         distance = factor(distance))
 

ggplot(Combined_frequency, aes(x=t, y=number_of_signals, size=distance, color=product)) +
  geom_point(alpha=0.5) +
  facet_wrap(~site) +
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))

ggplot(Combined_frequency, aes(x=date, y=number_of_signals, size=distance, color=product)) +
  geom_point(alpha=0.5) +
  facet_wrap(~site) +
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))

### Without time

ggplot(Combined_frequency, aes(x=distance, y= number_of_signals, color=product)) +
  geom_point(alpha=0.5) +
  facet_wrap(~site) +
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))

```

# Histogram

```{r}


combined_frequency_season <- Combined_frequency %>% 
  mutate(month = month(date, abbr = T, label = T),
         year = year(date)) %>% 
  mutate(season = ifelse(month %in% c("Jan", "Feb", "Mar"), "Summer",        
                           ifelse(month %in% c("Apr", "May", "Jun"), "Autumn",
                                ifelse(month %in% c("Jul", "Aug", "Sep"), "Winter",
                                       ifelse(month %in% c("Oct", "Nov", "Dec"), "Spring","Error")))))


CMC_dur <- read_excel("Data/CMC_dur.xlsx")

ggplot(data = CMC_dur, aes(x = duration)) +
  geom_histogram(aes(fill = product),  
                       position = "dodge", alpha=0.6) +
   labs(x = "Duration (Days)", y = "Number of upwelling signals", colour = "Temperature products") +
  facet_wrap(~site) +
theme_bw() 



ggplot(CMC_dur, aes(duration, colour = product)) +
  geom_freqpoly() +
  facet_wrap(~ site)
             
month_year_func <- function(df){
  month_year_func<- df %>% 
    mutate(month = month(date_start, abbr = T, label = T)) %>% 
    separate(date_start, into = c("year", "month_no")) %>% 
    unite(month_year, month,year)
    # Now need to get the year into a seperate column from the start date. So now date_start is split in the dataset
}

OISST_prep <- month_year_func(df = OISST_final)
CMC_prep <- month_year_func(df = CMC_final)
G1SST_prep <- month_year_func(df = G1SST_final)
MUR_prep <- month_year_func(df = MUR_final)

# Combined dataset
# Combined_frequency <- read_excel("Data/Combined_frequency.xlsx")
# save(Combined_frequency, file = "Data/Combined_frequency.RData")
```

# Plotting

```{r}
load("Data/Combined_frequency.RData")

Combined_frequency <- Combined_frequency %>%
  mutate(t = parse_date(date, format = "%b_%Y"),
         distance = factor(distance))

ggplot(data = Combined_frequency, aes(x = distance, y = number_of_signals)) +
  geom_boxplot(aes(colour = product)) +
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))+
  #geom_smooth(aes(colour = product), method = "lm") +
  #labs(x = NULL) +
  facet_wrap(~site) 


ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals, colour = product)) +
geom_point(aes(size = distance)) +
geom_smooth(method = "lm", size = 1.2)+
theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))

# Does not accurately show the months
# Plot need to include month_year, number of signals, distance and product ()

ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
  geom_line(aes(colour = product)) +
  geom_smooth(aes(colour = product), method = "lm") +
  labs(x = NULL) +
  facet_wrap(~site) #+
  # theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))


ggplot(data = Combined_frequency, aes(x = t, y = number_of_signals)) +
  geom_line(aes(colour = distance)) +
  geom_point(aes(colour = distance)) +
  geom_smooth(aes(colour = distance), method = "lm") +
  # labs(x = NULL) +
  facet_wrap(~site) #+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))
  
  ggplot(data = Combined_frequency, aes(x = date, y = number_of_signals)) +
  geom_line(aes(colour = distance)) +
  geom_point(aes(colour = distance)) +
  geom_smooth(aes(colour = distance), method = "lm") +
  # labs(x = NULL) +
  facet_wrap(~site) #+
  theme(axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5))

####

combined_frequency_split <- Combined_frequency %>% 
  separate(date, into = c("month", "year"))

# ggplot(data = combined_frequency_split, aes(x = year, y = number_of_signals)) +
#   geom_line(aes(colour = month)) +
#   geom_smooth(aes(colour = month), method = "lm") +
#   facet_wrap(~site)
```


### Looking at intensity

```{r}

filt_func <- function(df){
  intens <- df %>% 
  #mutate(year = year(date_start)) %>% 
  filter(year(date_start) %in% seq(2011, 2012))
}

filt_func2 <- function(df){
  intens1 <- df %>% 
  #mutate(year = year(date_start)) %>% 
  filter(year(date_start) %in% seq(2013, 2014))
}

OISST_filt_func <- filt_func(OISST_final)
OISST_filt_func2 <- filt_func2(OISST_final)
OISST_binding <- rbind(OISST_filt_func,OISST_filt_func2)# %>% 
 # select(site, product, intensity_mean)


MUR_filt_func <- filt_func(MUR_final)
MUR_filt_func2 <- filt_func2(MUR_final)
MUR_binding <- rbind(MUR_filt_func,MUR_filt_func2)# %>% 
 # select(site, product, intensity_mean)

G1SST_filt_func <- filt_func(G1SST_final)
G1SST_filt_func2 <- filt_func2(G1SST_final)
G1SST_binding <- rbind(G1SST_filt_func,G1SST_filt_func2) #%>% 
  #select(site, product, intensity_mean)


CMC_filt_func <- filt_func(CMC_final)
CMC_filt_func2 <- filt_func2(CMC_final)
CMC_binding <- rbind(CMC_filt_func,CMC_filt_func2) #%>% 
 # select(site, product, intensity_mean)

SACTN_upwell_base_ <- read_csv("Data/SACTN_upwell_base .csv") # The SACTN product column was added in 

SACTN_filt_func <- filt_func(SACTN_upwell_base_)
SACTN_filt_func2 <- filt_func2(SACTN_upwell_base_)
SACTN_binding <- rbind(SACTN_filt_func,SACTN_filt_func2)# %>% 
# write.csv(SACTN_binding,'SACTNfinal_data.csv')

Complete_mean_intensity <- rbind(CMC_binding,OISST_binding, G1SST_binding, MUR_binding) %>% 
  ungroup() %>% 
  select(-heading, - distance, -distance_km) # Here I remove distance, Keep if distance is needed

#write.csv(Complete_mean_intensity ,'Complete_mean_intensity.csv') # Copied this and then removedsome columnss

final_data <- rbind(Complete_mean_intensity, SACTN_binding)
# write.csv(final_data,'final_data.csv')

```

# PLotting

```{r}

signals_average <- read_csv("Data/signals_average.csv")

ggplot(data = signals_average, aes(x = no_of_signals, y = intensity_mean)) +
  geom_point(aes(colour = product), size = 3) +
  geom_smooth(method = "lm") +
  facet_wrap(~site) + 
  labs(x = "Number of signals ", y = "Average intensity (°C)") +
  scale_colour_Publication()+
  theme_Publication()



ggplot(data =final_intensity, aes(x = intensity_mean)) +
  geom_histogram(aes(fill = product),  
                       position = "dodge", alpha=0.6) +
  # labs(x = "Duration (Days)", y = "Number of upwelling signals", colour = "Temperature products") +
  facet_wrap(~site) +
```
# Loading in all the data created using the code bellow

```{r}
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
load("Data/site_pixels.RData") # 5 decimal places
load("Data/OISST.RData") # 2 decimal places
OISST <- BC_avhrr_only_v2_Document_Document 
rm(BC_avhrr_only_v2_Document_Document ); gc()
load("Data/CMC.RData") # 1decimal places
```



```{r}
# SACTN_US <- SACTN_US %>% 
#   dplyr::rename(in_situ_temp = temp)
# 
# # Visualising the data
# temp_plot <- function(df){
#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
#     geom_line(aes(group = site)) +
#     labs(x = "", y = "Temperature (°C)") +
#     theme(axis.text.x = element_text(angle = 45)) +
#     theme(legend.position = "top")
# }
# 
# SACTN_plot <- temp_plot(df = SACTN_US)
# SACTN_plot


# 25th quantile
```



# Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
# Is this upwelling event seen throughout the different distances from the coastline
# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets


```{r}
#glm to determine if there is a difference in the number of events recorded at different at every distance

load("Data/Count/OISST_count.RData")
load("Data/Count/G1SST_count.RData")
load("Data/Count/MUR_count.RData")
load("Data/Count/CMC_count.RData")
G1SST_count <- G1SST_count %>% 
  mutate(src = "G1SST")

combined <- rbind(CMC_count,G1SST_count,MUR_count,OISST_count)

options(scipen = 999)

anova_func <- function(df){
  sites_aov <- aov(distance ~ no_observation * src, data = df)
return(sites_aov)
}

combined <- anova_func(df = combined)
summary(combined)

m_coeff <- function(df){
  res <- lm(formula = val ~ date_peak, data = df)
  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
}
# Changes in upwelling metrics
lm_func <- function(df){
  upwell_lm <- df %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var, distance) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
}

OISST_lm <- lm_func(df = OISST_upwell_base)
CMC_lm <- lm_func(df = CMC_upwell_base)

SACTN_lm <- SACTN_upwell_base %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
```
  