---
title: "Remotely SST"
author: "Amieroh Abrahams"
date: "03 August 2019"
output: html_document
---

# Loading libraries

```{r}
library(tidyverse)
library(plyr)
library(lubridate)
library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(data.table)
library(heatwaveR)
library(viridis)
library(gridExtra)
library(fossil)
library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ggpubr)
library(ggpubr); theme_set(theme_pubr())# library(devtools)
library(FNN)
```

# Plots

```{r}
OISST %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
```

## MUR dataset

```{r}
load("~/Documents/Masters_2019/Upwelling_MCS/Data/SACTN_US.RData")

MUR_Lamberts_Bay <- read_csv("Data/MUR/MUR_Lamberts Bay_SST_timeseries_5nearest.csv")
MUR_Port_Nolloth <- read_csv("Data/MUR/MUR_Port Nolloth_SST_timeseries_5nearest.csv")
MUR_Saldanha_Bay <- read_csv("Data/MUR/MUR_Saldanha Bay_SST_timeseries_5nearest.csv")
MUR_Sea_Point <- read_csv("Data/MUR/MUR_Sea Point_SST_timeseries_5nearest.csv")
MUR_Hout_Bay<- read_csv("Data/MUR/MUR_Hout Bay_SST_timeseries_5nearest.csv")

MUR_SST <- rbind(MUR_Lamberts_Bay, MUR_Port_Nolloth,MUR_Saldanha_Bay, MUR_Hout_Bay, MUR_Sea_Point) %>%
  dplyr::rename(site = station)

MUR_SST$date <- (ymd(MUR_SST$date))
MUR_SST <- MUR_SST %>% 
  dplyr::rename(temp_MUR = nearest1) %>% 
  drop_na()

# save(MUR_SST, file = "Data/MUR_SST.RData")

match_func <- function(df){
  match <- SACTN_US  %>%  
  left_join(df, by = c("site", "date")) %>% 
  na.trim()
  return(match)
}

insitu_MUR <- match_func(df = MUR_SST) 
```

# Upwelling and MCS

```{r}

MUR_SST <- insitu_MUR %>% 
  select(date,temp_MUR, site) %>% 
  dplyr::rename(t = date) %>% 
  dplyr::rename(temp = temp_MUR)

filter_func <- function(df){
  filtered <- MUR_SST  %>%  
  filter(site == df) %>%
  select(-site) %>% 
  drop_na()
  return(filtered)
}

MUR_LB <- filter_func(df = "Lamberts Bay") 
MUR_SB <- filter_func(df = "Saldanha Bay") 
MUR_HB <- filter_func(df = "Hout Bay") 
MUR_PN <- filter_func(df = "Port Nolloth") 
MUR_SP <- filter_func(df = "Sea Point") 

# MCS

MCS_func <- function(df){
  res_clim <- ts2clm(df, climatologyPeriod = c("2002-06-01", "2015-12-30"),
                   pctile = 10)
  return(res_clim)
}

PN_MCS <- MCS_func(df = MUR_PN)
LB_MCS <- MCS_func(df = MUR_LB)
SB_MCS <- MCS_func(df = MUR_SB)
SP_MCS <- MCS_func(df = MUR_SP)

HB_MCS <- ts2clm(MUR_HB, climatologyPeriod = c("2002-06-01", "2005-10-19"), # Climatological period differs to the rest
                   pctile = 10)

## Events
MCS_event <- function(df){
  detect <- detect_event(df, coldSpells = TRUE)
  return(detect)
}

PN_MCS_events_MUR <- MCS_func(df = PN_MCS)
LB_MCS_events_MUR <- MCS_func(df = LB_MCS)
SB_MCS_events_MUR <- MCS_func(df = SB_MCS)
SP_MCS_events_MUR <- MCS_func(df = SP_MCS)

HB_MCS_events_MUR <- detect_event(HB_MCS, coldSpells = TRUE) 
```

## Upwelling events - 30th percentile used as a threshold to detect upwellling events. 

```{r}
upwell_func <- function(df){
  res_clim <- ts2clm(df, climatologyPeriod = c("2002-06-01", "2015-12-30"),
                   pctile = 30)
  return(res_clim)
}

PN_upwell <- upwell_func(df = MUR_PN)
LB_upwell <- upwell_func(df = MUR_LB)
SB_upwell <- upwell_func(df = MUR_SB)
SP_upwell <- upwell_func(df = MUR_SP)

HB_upwell <- ts2clm(MUR_HB, climatologyPeriod = c("2002-06-01", "2005-10-19"), # Climatological period differs to the rest
                   pctile = 10)

## Events
upwell_event <- function(df){
  detect <- detect_event(df, minDuration = 2, coldSpells = TRUE)
  return(detect)
}

PN_upwell_events_MUR <- MCS_func(df = PN_upwell)
LB_upwell_events_MUR <- MCS_func(df = LB_upwell)
SB_upwell_events_MUR <- MCS_func(df = SB_upwell)
SP_upwell_events_MUR <- MCS_func(df = SP_upwell)

HB_upwell_events_MUR <- detect_event(HB_upwell, minDuration = 2, coldSpells = TRUE)
```

## CMC dataset

```{r}
Lamberts_Bay <- read_csv("Data/CMC/Lamberts Bay_SST_timeseries_5nearest.csv")
Port_Nolloth <- read_csv("Data/CMC/Port Nolloth_SST_timeseries_5nearest.csv")
Saldanha_Bay <- read_csv("Data/CMC/Saldanha Bay_SST_timeseries_5nearest.csv")
Sea_Point<- read_csv("Data/CMC/Sea Point_SST_timeseries_5nearest.csv")
Hout_Bay <- read_csv("Data/CMC/Hout Bay_SST_timeseries_5nearest.csv")

CMC <- rbind(Lamberts_Bay, Hout_Bay, Port_Nolloth, Saldanha_Bay, Sea_Point) %>%
  dplyr::rename(site = station)

CMC$date <- (ymd(CMC$date))
CMC <- CMC %>% 
  drop_na() %>% 
  dplyr::rename(temp_CMC = nearest1)

# save(CMC, file = "Data/CMC_SST.RData")

## Matching the Insitu data with the CMC SST data
insitu_CMC <- match_func(df = CMC) 

CMC_SST <- insitu_CMC %>% 
  select(date,temp_CMC, site) %>% 
  dplyr::rename(t = date) %>% 
  dplyr::rename(temp = temp_CMC)

filter_func <- function(df){
  filtered <- CMC_SST  %>%  
  filter(site == df) %>%
  select(-site) %>% 
  drop_na()
  return(filtered)
}

CMC_LB <- filter_func(df = "Lamberts Bay") 
CMC_SB <- filter_func(df = "Saldanha Bay") 
CMC_HB <- filter_func(df = "Hout Bay") 
CMC_PN <- filter_func(df = "Port Nolloth") 
CMC_SP <- filter_func(df = "Sea Point") 

# MCS

MCS_func <- function(df){
  res_clim <- ts2clm(df, climatologyPeriod = c("1991-09-01", "2015-12-30"),
                   pctile = 10)
  return(res_clim)
}

PN_MCS <- MCS_func(df = CMC_PN)
LB_MCS <- MCS_func(df = CMC_LB)
SB_MCS <- MCS_func(df = CMC_SB)
SP_MCS <- MCS_func(df = CMC_SP)

HB_MCS <- ts2clm(CMC_HB, climatologyPeriod = c("2002-06-01", "2005-10-19"), # Climatological period differs to the rest
                   pctile = 10)

## Events
MCS_event <- function(df){
  detect <- detect_event(df, coldSpells = TRUE)
  return(detect)
}

PN_MCS_events_CMC <- MCS_func(df = PN_MCS)
LB_MCS_events_CMC <- MCS_func(df = LB_MCS)
SB_MCS_events_CMC <- MCS_func(df = SB_MCS)
SP_MCS_events_CMC <- MCS_func(df = SP_MCS)

HB_MCS_events_CMC <- detect_event(HB_MCS, coldSpells = TRUE) 
```

## Upwelling events - 30th percentile used as a threshold to detect upwellling events. 

```{r}
upwell_func <- function(df){
  res_clim <- ts2clm(df, climatologyPeriod = c("1991-09-01", "2015-12-30"),
                   pctile = 30)
  return(res_clim)
}

PN_upwell <- upwell_func(df = CMC_PN)
LB_upwell <- upwell_func(df = CMC_LB)
SB_upwell <- upwell_func(df = CMC_SB)
SP_upwell <- upwell_func(df = CMC_SP)

HB_upwell <- ts2clm(CMC_HB, climatologyPeriod = c("2002-06-01", "2005-10-19"), # Climatological period differs to the rest
                   pctile = 10)

## Events
upwell_event <- function(df){
  detect <- detect_event(df, minDuration = 2, coldSpells = TRUE)
  return(detect)
}

PN_upwell_events_CMC <- MCS_func(df = PN_upwell)
LB_upwell_events_CMC <- MCS_func(df = LB_upwell)
SB_upwell_events_CMC <- MCS_func(df = SB_upwell)
SP_upwell_events_CMC <- MCS_func(df = SP_upwell)

HB_upwell_events_CMC <- detect_event(HB_upwell, minDuration = 2, coldSpells = TRUE)
```

## OISST

```{r}
OISSTDir <- "~/Documents/Masters_2019/Data_SST"
OISST <- fread(paste0(OISSTDir, "/csvavhrr-only-v2-19810901-20180630.csv"),
            col.names = c("lon", "lat", "temp", "date"))

# Visualise the data
# To explore the data I visualise the min temperatures along the South african coastline.

OISST %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = temp))
```

## The nearest SST pixels
Now we apply the FNN (Fast Nearest Neighbor) package to determine the nearesr SST pixel to the insitu collected sites. 
```{r}
load("~/Documents/Masters_2019/Upwelling_MCS/Data/site_list_sub.Rdata")

unique_pixel <- OISST %>%
  select(lon, lat) %>%
  unique()

# Select nearest 1 pixels (k = 1)
# Here I use knnx to find the closes 1 pixels to the insitu sites
match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                      query = as.matrix(site_list_sub[,5:6]), k = 1)

pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = site_list_sub$site)

# Subsetting the OISST data to match the upwelling sites within the in situ collected temperature data
OISST_match <- OISST %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)
# save(OISST_match, file = "Data/OISST_match.RData")

OISST_sites <- OISST_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>%
  dplyr::rename(temp_OISST =temp)

OISST_sites <- OISST_sites %>% 
  dplyr::rename(temp = temp_OISST) %>%
  dplyr::mutate(date = as.Date(date)) %>% 
  drop_na()
# save(OISST_sites, file = "Data/OISST_sites.RData")
```

```{r}
load("~/Documents/Masters_2019/Upwelling_MCS/Data/OISST_sites.RData")
## Matching the Insitu data with the OISST SST data
insitu_OISST <- match_func(df = OISST_sites)
```





